### 2021-11-19 mk2/manuscript.rmd
updated manuscript

---
title: "Erroneous use of Venn diagram leads to artifacts in transcriptomic analyses"
author: "January Weiner^1,†^, Benedikt Obermayer^1^ and Dieter Beule^1^"
date: "`r Sys.Date()`"
outputoff:
  html_document
output:
  word_document:
    reference_docx: templates/template.docx
  pdf_document:
    latex_engine: xelatex
bibliography: manuscript.bib
link-citations: true
XXcsl: templates/nature-medicine.csl
---

^1^ Berlin Institute of Health at Charité – Universitätsmedizin Berlin, Core Unit Bioinformatics, Charitéplatz 1, 10117 Berlin, Germany

^†^ To whom the correspondence should be addressed

## Abstract

Venn diagrams are frequently used to illustrate the results of
differential expression analysis and further
downstream functional analysis. Here we show that both, the use
of Venn diagrams to find genes which are thought to be specific for a
certain comparison, as well as gene set enrichment analysis applied to
such subsets is a fallacy. Moreover, the combination of incorrect statistical reasoning
and gene set enrichment analysis may result in particularly misleading artifacts.

```{r,echo=FALSE}
## Set default options for the knitr RMD processing
knitr::opts_chunk$set(echo=FALSE,warning=FALSE,message=FALSE,fig.width=5,fig.height=5,cache=FALSE,autodep=TRUE, results="hide")
```

```{r libraries,cache=FALSE}
library(tidyverse)
library(limma)
library(GEOquery)
library(DESeq2)
## install bioshmods from github.com/bihealth/bioshmods
library(bioshmods)
library(tmod)
library(cowplot)
library(stringr)
library(ggvenn)
library(ggplot2)
library(ggpval)
library(eulerr)
library(Biobase)
library(org.Hs.eg.db)
theme_set(theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()))
source("functions.R")
```

```{r setup}
padj_thr <- 0.05
lfc_thr  <- 1
```

## Introduction: Venn diagrams are a showcase of erroneous interpretation of interactions

Venn diagrams (VDs) are commonly used to visualize high throughput data
such as transcriptomic profiles. Frequently, a Venn diagram serves the
purpose of comparing the results from two experiments or two comparisons.
For example, a VD may illustrate up- or down-regulated genes in two strains
of mice upon infection, showing which genes are regulated in both strains,
and which are regulated only in one of them.  The VD serves
as a basis for the statement that the regulation of certain genes is
"specific" for one strain or another. We argue that not only is this
inference incorrect, but it may also lead to misleading – although
appealing – artifacts when combined with downstream analyses.

Consider a gene for which the expression has been analysed in four groups:
two different mouse strains (wildtype, WT and knockout, KO) and two different experimental conditions
(naive versus infected). We find that the gene expression significantly
differs between infected and naive KO animals, but that there is no
significant difference in the WT strain. Such a gene may be incorrectly
considered as "specific" to KO, and will be accordingly entered in a Venn
diagram.

However, this is a common fallacy [@nieuwenhuis2011erroneous], since lack
of statistical significance is not the same as testing for lack of
difference. In other words, the fact that we failed to detect a significant
difference in the WT does not mean that the difference is absent and
itself significantly different from the difference in the KO. This "difference of
differences" is known is statistics as interaction [@blalock1965theory].
Here, it is the interaction between strain and treatment.  In fact, the
obtained p-values might be just over the assumed p-value threshold in one, but just
under it in the other strain (e.g. 0.009 vs 0.011), with the difference in
both strains being essentially the same. The correct analysis is to test
the significance of interaction using an ANOVA model. Notably, in 2011
Nieuwenhuis et al. found that in over 150 papers in top neuroscience
journals where the authors could make this particular statistical error,
half of them did.

Using VDs to show genes "specific" for a condition amounts to
counting the times a comparison for a gene was statistically significant in
one condition, but not significant in another.
More disturbingly, VDs are a visual illustration of a procedure
that exacerbates this problem by applying a downstream analysis to
supposedly specific sets of genes. Several cells in a Venn diagram contain
numbers corresponding to features (such as genes), for which there was a significant test result in
one, but not in another comparison. The sets of genes in a Venn diagram cell may
then be subsequently analysed to test whether they share a particular
characteristics.  For example, gene set enrichment analysis may be used to
interpret the biological function of genes that are considered "specific" to one
condition. It turns out that due to the
fallacious nature of this procedure, it is likely to produce results that
appear to make sense in the given biological context.

In this paper, we illustrate this rather simple statistical statement with
a real world data set, demonstrating how choosing this sort of approach
results in apparently sound gene set enrichment results which are, in fact,
artifacts. Next, we dissect the underlying mechanism of how these artifacts
are generated.  Finally, we discuss alternative approaches.

## Results

### Transcriptomic changes due to Sars-Cov-2 infection

```{r}
n_stratum <- 10 # 2 strata per condition 
```

Consider two group of patients, G1 and G2. Each group contains 
`r n_stratum * 4`
individuals. In both groups, there is an equal number of healthy
individuals (labeled "Ctrl" on figures below) 
or patients infected with Sars-Cov-2 (labeled "SC2"). Our aim is to understand
the differences between G1 and G2 in the response to infection. For
example, we ask which genes or pathways are specifically upregulated by SC2
infection in G1 as compared to G2, and vice versa. In the following, we
used the data set GSE156063 [@mick2020upper] in two approaches (an
incorrect and the correct one) which arrive at opposite conclusions.

```{r pheno_data_download,eval=FALSE}
## this sometimes fails. The RDS of the object is provided
Sys.setenv(VROOM_CONNECTION_SIZE=8*131072)

## with getGEO, we can only get the phenoData
geo <- getGEO("GSE156063")[[1]]
saveRDS(geo, file="GSE156063.geo")
```

```{r pheno_data_cleanup}
geo <- readRDS("GSE156063.geo")
covar <- as(phenoData(geo), "data.frame")

## remove columns with only one value
boring <- map_lgl(covar, ~ length(unique(.x)) == 1)
covar <- covar[ , !boring ]

## clean up
covar <- covar %>% 
  dplyr::rename(gender = "gender:ch1") %>%
  mutate(disease = gsub(" .*", "", .data[["disease state:ch1"]])) %>%
  mutate(label = description) %>%
  mutate(group = disease) %>%
  arrange(description) %>%
  dplyr::select(all_of(c("title", "label", "gender", "disease", "group")))
```

```{r featuredata_download}
## the counts must be downloaded from GEO separately.
if(!file.exists("GSE156063_swab_gene_counts.csv.gz")) {
  download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE156nnn/GSE156063/suppl/GSE156063_swab_gene_counts.csv.gz",
                "GSE156063_swab_gene_counts.csv.gz")
}

counts <- read_csv("GSE156063_swab_gene_counts.csv.gz")
.tmp <- counts[[1]]
counts <- as.matrix(counts[,-1])
rownames(counts) <- .tmp
#counts <- counts[ , covar$description ]
counts <- counts[ , covar$label ]
lcpm   <- edgeR::cpm(counts, log=TRUE)
#stopifnot(all(colnames(counts) == covar$description))
stopifnot(all(colnames(counts) == covar$label))

annot <- data.frame(ENSEMBL = rownames(counts))

.tmp <- mapIds(org.Hs.eg.db, annot$ENSEMBL, column=c("ENTREZID"), keytype="ENSEMBL")
annot$ENTREZID <- .tmp[ match(annot$ENSEMBL, names(.tmp)) ]

.tmp <- mapIds(org.Hs.eg.db, annot$ENSEMBL, column=c("SYMBOL"), keytype="ENSEMBL")
annot$SYMBOL <- .tmp[ match(annot$ENSEMBL, names(.tmp)) ]

.tmp <- mapIds(org.Hs.eg.db, annot$ENSEMBL, column=c("GENENAME"), keytype="ENSEMBL")
annot$GENENAME <- .tmp[ match(annot$ENSEMBL, names(.tmp)) ]
```

```{r Preparation_of_the_covariates,cache=TRUE}
sel <- covar$group %in% c("no", "SC2")
counts <- counts[ , sel ]
covar  <- covar[ sel, ]
covar$group <- as.character(covar$group)

set.seed(0123)

#covar$disease <- covar$group

g1 <- covar %>% mutate(n=1:n()) %>%
  group_by(gender, disease) %>% slice_sample(n=n_stratum) %>% pull(n)
g2 <- covar %>% mutate(n=1:n()) %>% filter(!n %in% g1) %>%
  group_by(gender, disease) %>% slice_sample(n=n_stratum) %>% pull(n)

#g1 <- sample(1:nrow(covar), 30)
#g2 <- sample(setdiff(1:nrow(covar), g1), 30)

covar$group <- NA
covar$group[g1] <- "G1"
covar$group[g2] <- "G2"

covar$group.disease <- paste0(covar$group, '_', covar$disease)

sel <- c(g1, g2)

ds2 <- DESeqDataSetFromMatrix(counts[,sel], colData=covar[sel, ],
  design=~ 0 + group.disease )
```

```{r DESeq2,cache=TRUE}
## DESeq2 calculations

ds2 <- DESeq(ds2)
saveRDS(ds2, file="ds2_cached.rds")
```

```{r DE_analysis}
library(tidyverse)
library(tmod)
res <- list()
res$g1 <- results(ds2, contrast=c(-1, 1, 0, 0))
res$g2 <- results(ds2, contrast=c(0, 0, -1, 1))

res <- map(res, ~ .x %>% as.data.frame() %>%
  rownames_to_column("ENSEMBL") %>% 
  left_join(annot, by="ENSEMBL") %>%
  mutate(DEG=!is.na(padj) & abs(log2FoldChange) > lfc_thr & padj < padj_thr))

res.merged <- merge(res$g1, res$g2, by=c("ENSEMBL", "SYMBOL", "ENTREZID", "GENENAME"), suffixes=c(".g1", ".g2"))
```

First, we have performed differential gene expression analysis for each of
the groups G1 and G2 separately using standard bioinformatic tools. For each comparison, we defined
genes with significantly different gene expression 
(differentially expressed genes, DEGs) 
by using an FDR threshold of 
`r padj_thr`
and absolute log~2~ fold change (LFC) threshold of
`r lfc_thr`. There were
`r sum(res$g1$DEG)` 
DE genes in the G1 group, and
`r sum(res$g2$DEG)`
in the G2 group. In total,
`r sum(res.merged$DEG.g1 & res.merged$DEG.g2)` 
DEGs were common for G1 and G2, 
`r sum(res.merged$DEG.g1 & !res.merged$DEG.g2)` 
DEGs were significant in G1 only ("specific" for G1), and 
`r sum(!res.merged$DEG.g1 & res.merged$DEG.g2)` 
were significant in G2 only ("specific" for G2; see Fig. 1, panel *A*).
A naive interpretation of these results implies that there is a substantial
difference between these two groups of individuals, as evidenced by a small
overlap in commonly regulated genes, while the majority of DEGs is
significant in one comparison only.

```{r GO_db_prepare,cache=TRUE}
library(msigdbr)
min_mod_size <- 10
max_mod_size <- 50
msig_go_bp <- msigdbr(subcategory="GO:BP")
mset <- makeTmodFromDataFrame(msig_go_bp, 
                              feature_col="human_gene_symbol", 
                              module_col="gs_exact_source", 
                              title_col="gs_name")
mset$MODULES$N <- map_int(mset$MODULES$ID, ~ length(mset$MODULES2GENES[[.x]]))
mset <- mset[ mset$MODULES$N <= max_mod_size & 
              mset$MODULES$N >= min_mod_size ]
```

```{r GSEA}
common <- res.merged %>% filter(DEG.g1 & DEG.g2) %>% pull(SYMBOL)
gsea_res <- map(res, ~ {
  fg <- .x %>% filter(padj < padj_thr & abs(log2FoldChange) > lfc_thr)
  fg <- setdiff(fg$SYMBOL, common) # only "specific" genes
  tmodHGtest(fg=fg, bg=.x$SYMBOL, mset=mset)
})
gsea_res_full <- map(res, ~ {
  fg <- .x %>% filter(padj < padj_thr & abs(log2FoldChange) > lfc_thr)
  fg <- setdiff(fg$SYMBOL, common) # only "specific" genes
  tmodHGtest(fg=fg, bg=.x$SYMBOL, mset=mset, qval = Inf)
})
gsea_res.merged <- merge(gsea_res$g1, gsea_res$g2, by=c("ID", "Title"), suffixes=c(".g1", ".g2"), all=T)
```

To understand which
pathways are upregulated in each of the two groups, we used a standard generation I
gene set enrichment analysis (GSEA) – a hypergeometric test – on the DEGs in each
group. Gene sets for GSEA were taken from the Gene Ontology (GO) database.
Gene sets with more than `r max_mod_size` or 
fewer than `r min_mod_size` genes were removed.
For each group, we have
selected only genes which are DEGs in that group, but not the other,
mimicking a naive approach for finding "specifically" regulated pathways.
Here, a similar picture
emerged. Just
`r nrow(gsea_res$g1)` gene sets were significantly enriched in G1, and
`r nrow(gsea_res$g2)` gene sets were significantly enriched in G2. Again,
both the Venn diagram (Fig. 1, B) and the results of
enrichments (Fig. 1, C and D) suggest that there is a
fundamental difference between the groups, and that the groups have little
in common in their response to the virus.

Importantly, the different GO terms enriched in the two groups were related
to infectious diseases, and may tempt to speculate about the underlying
biological differences between these two groups (e.g.  significance of
Toll like receptor 4 pathway in G1, but not G2; and, vice versa, significance of
response to interleukin 7 in G2, but not in G1).

**Fig. 1. Results of differential gene expression analysis
and gene set enrichment analysis using an incorrect Venn diagram driven approach.**
**A**, Venn diagram showing numbers of differentially expressed genes (DEG)
in each of the two groups, G1 and G2; **B**, Venn diagram showing numbers
of significantly enriched GO terms in each of the two groups; **C** results
of gene set enrichment analysis for genes "specific" to group G1; **D**, results of gene set
enrichment analysis for genes "specific" to group G2 (only top 10 terms are shown).

```{r fig1, fig.width=15,fig.height=5.5,dpi=300,dev="png"}
p1 <- ggvenn(list(G1=res$g1 %>% filter(DEG) %>% pull(ENSEMBL),
                         G2=res$g2 %>% filter(DEG) %>% pull(ENSEMBL)),
             show_percentage=FALSE) +
      ggtitle("Differentially expressed genes")
p2 <- ggvenn(list(G1=gsea_res$g1 %>% pull(ID),
                         G2=gsea_res$g2 %>% pull(ID)),
              show_percentage=FALSE) +
      ggtitle("Enriched GO terms")

col1 <- plot_grid(p1, p2, labels="AUTO", nrow=2)

p3 <- plot_gsea(gsea_res$g1 %>% arrange(P.Value) %>% dplyr::slice(1:10)) + ggtitle("G1") +
  scale_x_discrete(labels=function(x) str_wrap(x, width=25))
#p3 <- plot_grid(p3, NULL, ncol=1, rel_heights=c(2, 1))
p4 <- plot_gsea(gsea_res$g2 %>% arrange(P.Value) %>% dplyr::slice(1:10)) + ggtitle("G2") +
  scale_x_discrete(labels=function(x) str_wrap(x, width=25))

plot_grid(plotlist = list(col1, p3, p4), labels=c('', 'C', 'D'), nrow=1)
```

The groups G1 and G2 were randomly sampled from the same data set.
In fact, repeated re-sampling always results in some genes being found to
be significantly different in one group, but not the other, despite the
fact that one does not expect any major differences between sets of
individuals randomly drawn from one population. Thus, the conclusions drawn
from a Venn diagram-driven gene set enrichment analysis are artefactual. Closer
inspection of genes which belong to DEG in one group, but not the other
reveals the underlying statistical fallacy (Fig. 2, *A* – *D*), that is, that difference
between significant and non-significant is, in itself, not statistically
significant [@gelman2006difference]. This does not necessarily mean that
there are no differences at all between these two groups, but that lack of
significance in one group and significance in the other group does not
correctly identify differences between groups.

```{r interaction}
res$int <- results(ds2, contrast=c(-1, 1, 1, -1)) %>%
  as.data.frame() %>%
  rownames_to_column("ENSEMBL") %>% 
  left_join(annot, by="ENSEMBL") %>%
  mutate(DEG=!is.na(padj) & abs(log2FoldChange) > lfc_thr & padj < 0.05)

fg <- res$int %>% filter(padj < padj_thr & abs(log2FoldChange) > lfc_thr)
if(nrow(fg) > 0) {
  gsea_res$int <-  tmodHGtest(fg=fg$SYMBOL, bg=res$int$SYMBOL, mset=mset)
}

tmod_res <- list()
tmod_res$int <- tmodCERNOtest(res$int %>% arrange(pvalue) %>% pull(SYMBOL),
                              mset=mset)
```

```{r correlation_coefs}
## calculate correlation coefficients for various groups of genes
## all correlations are significant

cors <- list()
cors$all <- with(res.merged, 
                cor(log2FoldChange.g1, log2FoldChange.g2))
cors$sign <- with(res.merged %>% filter(DEG.g1 | DEG.g2), 
                cor(log2FoldChange.g1, log2FoldChange.g2))
cors$g1 <- with(res.merged %>% filter(DEG.g1 & !DEG.g2), 
                cor(log2FoldChange.g1, log2FoldChange.g2))
cors$g2 <- with(res.merged %>% filter(!DEG.g1 & DEG.g2), 
                cor(log2FoldChange.g1, log2FoldChange.g2))

```

To find genes which are differentially regulated in the two groups, the
correct statistical approach is to calculate interaction between groups
(G1, G2) and disease status (no disease vs. COVID). While it may be argued
that a test for interaction has lower power than a test for a simple
contrast, no 
genes show a significant interaction even at FDR < 0.1. In fact, this is not
surprising. The log~2~ fold changes for comparisons withing G1 and G2 are
strongly correlated (Fig. 2, *E*). For all significant genes, the Pearson correlation
coefficient is
`r format.pval(cors$sign, digits=2)`, 
while for genes
exclusively significant in G1 or G2 (genes "specific" to G1 or G2), it is
`r format.pval(cors$g1, digits=2)` 
and
`r format.pval(cors$g2, digits=2)`, 
respectively. Thus, genes which are significant in one, but not in the
other comparison often have have similar log~2~ fold changes (e.g. Fig.
2, *A*, *C* and *D*).

**Fig. 2. Genes which are significant in one comparison, but not
the other do not show a statistically significant interaction.** **A** -
**D**, examples of genes which are DEG in one group, but are not
significantly different in the other group. "Ctrl", healthy individuals; 
"SC2", Sars-Cov-2 infected patients. Values above the plot indicate FDR (p-values
corrected for multiple testing). **E**, correlation between log~2~ fold
changes in G1 and G2. Color indicates genes which are are significant in
one, but not significant in the other comparison; red indicates genes significant in G1, 
blue indicates genes significant in G2. The overall Pearson correlation
coefficient between log~2~ fold changes is 
`r with(res.merged, format.pval(cor(log2FoldChange.g1, log2FoldChange.g2), digits=2))`.

```{r fig2, fig.width=10, fig.height=5.5, dpi=300,dev="png"}
#id    <- gsub(" ", "_", toupper("response to interferon beta"))
#gstitle    <- "GOBP_TOLL_LIKE_RECEPTOR_4_SIGNALING_PATHWAY"
gstitle    <- "GOBP_INTERLEUKIN_7_MEDIATED_SIGNALING_PATHWAY"
gsid  <- gsea_res$g2 %>% filter(grepl(gstitle, Title)) %>% pull(ID)
gsgenes <- mset$MODULES2GENES[[gsid]]
sel   <- res.merged %>% filter(SYMBOL %in% gsgenes) %>% arrange(pvalue.g2) %>% 
  filter(DEG.g2 & !DEG.g1) %>%
  dplyr::slice(1:4) %>% pull(ENSEMBL)

sample_sel <- c(g1, g2)

group_lab <- gsub("_no", "/Ctrl", gsub("_SC2", "/SC2", covar$group.disease))

pees <- map(sel, ~ {
  s1 <- res$g1 %>% dplyr::slice(match(.x, ENSEMBL)) %>% pull(padj)
  s2 <- res$g2 %>% dplyr::slice(match(.x, ENSEMBL)) %>% pull(padj)
  id <- res$g1 %>% dplyr::slice(match(.x, ENSEMBL)) %>% pull(SYMBOL)

  ggplot_gene(lcpm[.x, sample_sel ], 
              group_lab[ sample_sel ],
              annotation=format.pval(c(s1, s2), digits=2), textsize=8,
              pval_text_adj=.5) +
            ggtitle(id)
})

part1 <- plot_grid(plotlist=pees, labels="AUTO", ncol=2)

tmp <- res.merged %>% mutate(specific= 
                             ifelse(DEG.g1 & !DEG.g2,
                                    "G1", 
                                    ifelse(!DEG.g1 & DEG.g2, "G2", "NS"))) %>%
                      mutate(tlr4=SYMBOL %in% gsgenes)

part2 <- ggplot(tmp, aes(x=log2FoldChange.g1, 
                                y=log2FoldChange.g2, 
                                color=specific)) + 
         scale_color_manual(values=c(G1="red", G2="blue", NS="#666666")) +
         geom_point(alpha=.5) + xlab("G1") + ylab("G2") +
         geom_hline(yintercept=0, color="grey") +
         geom_vline(xintercept=0, color="grey") +
         geom_abline(slope=1, intercept=0, color="grey") +
         guides(color=FALSE)

plot_grid(part1, part2, rel_widths=c(2, 2), labels=c("", "E"))
```
Consequently, it is not possible to calculate gene set enrichment for the
interaction using a
hypergeometric test, as there are no DEGs for the interaction contrast.
Gene set enrichment using a second generation algorithm (CERNO), relying on the
ordering of genes according to their raw p-values from the interaction
contrast rather than selecting a
set of DEGs [@zyla2019gene], does not show any significant enrichment.

### Artifacts arise because of false negatives

It is worth noting that in the gene set enrichment analysis of the genes
"specific" for a given comparison – i.e., genes which are significant in
that comparison, but not significant in others – we have observed a number
of terms associated with immune response. It is a crucial point of this
manuscript to note that the spurious enrichments not only show significant
p-values, but also that the terms or pathways which appear in them are
relevant to the research hypothesis being tested. Below, we will show why
these terms (rather than random terms which have no obvious relevance for
an infectious disease) appear in the results.

To understand how gene set enrichment actually gives significant
results in such randomly generated gene sets, despite absence of genes with
significant interaction, it is first necessary to consider the definition
of a differentially expressed gene in this context. More often than not,
DEGs are defined by a threshold in p-value adjusted for multiple testing,
possibly combined with a threshold in log~2~ fold change. The commonly used
Benjamini-Hochberg procedure [@benjamini1995controlling] ensures that among
genes for which the adjusted p~adj~ < 0.05 there are at most 5% false
positives.

This way, we can exert control over the false positive rate (FPR, type I
errors), keeping it at a relatively low level. However, we do not control
the false negative rate (FNR, type II errors).  In a powerful statistical
test (such as a t-test), the test power in a typical application will
rarely achieve more than 80%. For example, even for large
effects (Cohen's d > 0.8) and type I error rate of 0.05, a t-test only achieves
80% power with at least 
`r floor(pwr.t.test(power=.8, d=0.8)$n)`
samples per group. For small effects (Cohen's d > 0.2), the required number
of samples is at least
`r floor(pwr.t.test(power=.8, d=0.2)$n)`
per group.
Even assuming a test power of 80%, the FNR is 20%. Clearly, false negatives (FNs) occur
at much higher rates than false positives (FPs). In case of high throughput
data sets, where the FPR is controlled by Bejnamini-Hochberg procedure or a
similar technique, the FNR may be even as high as 80% [@white2019beyond].

```{r full_dataset}
ds2_full <- DESeqDataSetFromMatrix(counts, colData=covar, design=~ disease )
ds2_full<- DESeq(ds2_full)
res$full <- results(ds2_full, name="disease_SC2_vs_no") %>% as.data.frame() %>%
  rownames_to_column("ENSEMBL") %>% 
  left_join(annot, by="ENSEMBL") %>%
  mutate(DEG=!is.na(padj) & abs(log2FoldChange) > lfc_thr & padj < padj_thr)
```

```{r g_specific}
g1_spec <- res.merged %>% filter(DEG.g1 & !DEG.g2) %>% pull(ENSEMBL)
g2_spec <- res.merged %>% filter(!DEG.g1 & DEG.g2) %>% pull(ENSEMBL)
```

These FNs will occur
at a much higher rate within the sets of DEGs defined by the
non-overlaping areas of the VDs, that is DEGs considered to be "specific"
for one group or other in a naive approach.
To illustrate this phenomenon, we have analysed the full data set from which G1 and G2
were drawn (Fig. 3), comparing the 
`r sum(covar$disease == "no") `
healthy controls to 
`r sum(covar$disease == "SC2") `
COVID-19 patients. Of the
`r length(g1_spec)`
genes significant in G1, but not in G2, 
`r .n <- res$full %>% filter(DEG & ENSEMBL %in% g1_spec) %>% nrow; .n`
(`r sprintf("%.0f%%", .n/length(g1_spec) * 100)`) 
are significant in the full data set; 
of the
`r length(g2_spec)`
genes significant in G2, but not in G1, 
`r .n <- res$full %>% filter(DEG & ENSEMBL %in% g2_spec) %>% nrow; .n`
(`r sprintf("%.0f%%", .n/length(g2_spec) * 100)`) 
are significant in the full data set.
Given that G1 and G2 were sampled from the total population, and since the
FDR was set to `r padj_thr`, we can assume that at least between a third and a half of
the genes that appeared to be "specific" in the initial analysis were, in
fact, false negatives in one of the comparisons.

In other words, a substantial fraction of the "specific" genes are genes
that are in reality differentially expressed in both groups alike.
Therefore, if one is to perform a gene set enrichment analysis on one of
these "specific" groups of genes, then the enriched functions will be
related to the pathways and processes up- or down-regulated in both groups
due to the common factor (in this example, the COVID-19 disease), but
which are not related to differences between the two groups.

**Fig. 3. Area-proportional Venn diagram showing overlaps in DEGs between
G1, G2 and the full data set.** The majority of genes which have been labeled
as DEGs in only one of the groups G1, G2 are DEGs when all data were
analysed.

```{r fig3,dpi=300,dev="png"}
tmp <- merge(res.merged, res$full, by="ENSEMBL")
tmp <- tmp[ , c("DEG.g1", "DEG.g2", "DEG") ]
colnames(tmp) <- c("G1", "G2", "Full dataset")
plot(euler(tmp), quantities=list(type=c("counts", "percent")))
```

## Discussion

```{r lit_surves}
links <- system2("grep", args=c("https", "literature_survey.md"), stdout=TRUE)
n_tot <- length(links)
n_vd_only <- length(grep("vd only", links))
```

In an informal survey of articles from the journal "Scientific reports"
(chosen for its broad scope, large number of published articles and open
access), of the 238 articles from 2021 which used the term "venn diagram" and
"differential expression", at least 
`r n_tot`
were using Venn diagrams to compare
statistical significance with lack thereof by referring to "unique",
"specific", "solely regulated" or "exclusive" DEGs. Out of these, at least 
`r n_vd_only`
coupled the VDs with some form of gene set enrichment analysis on the
set of supposedly "specific" DEGs.

Drawing conclusions from comparing significance with lack thereof is a
common statistical fallacy [@gelman2006difference]. Just as absence of evidence is not evidence of
absence, the failure to reject the null hypothesis does not consitute the
same level of evidence as rejecting it. However, when such an incorrect
analysis is combined with downstream functional analysis – i.e., gene set
enrichments of genes "specific" to one or another comparison – the
resulting pathways or gene ontologies are misleadingly relevant, yielding
gene sets associated with immune answer for research hypotheses involving
an infectious disease, or cancer pathways if the underlying research
hypothesis involved cancer treatment. Such results may be hard to resist
for an involved researcher, especially if the correct analysis of
interactions does not show any significant differences.

That is not to say that VDs are not a useful tool, even in the context of
transcriptomics and gene set enrichments, if used correctly. For example, analysis of an
intersection of DEGs (i.e., by considering genes from the overlap in a VD)
is not an incorrect procedure.  Genes in the overlapping part of a VD are
significant in both (or all) comparisons, hence no comparison between
significance and non-significance is made.

As an alternative to Venn diagrams two approaches can be considered.
Firstly, as shown above, a test for interaction can reveal genes for which
the impact of treatment significantly differ between the groups. As a mean
of visualization, we recommend plotting the log~2~ fold changes in one
comparison versus the log~2~ fold changes in the other comparison. In
addition, a concordance / discordance analysis may provide additional
insight in genes which are regulated either in the same direction or in the
opposite directions in both groups [@domaszewska2017concordant].

While the use of Venn diagrams in combination with the incorrect gene set
enrichment analysis is common, it is not the only way of achieving
artifactual and incorrect conclusions by means of comparing significant
with non-significant results in transcriptomics and other high-throughput
applications. Our literature survey showed that at least two other ways are
common. Firstly, the direct comparison of gene set enrichment
results: that is, drawing conclusions from the fact that a gene set
enrichment result was significant in one comparison only.
Second, while Venn diagrams are often used to illustrate the numbers of
"specific" DEGs and so present a mean to find examples of this fallacy in
scientific literature, researches often test for enrichment these
"specific" genes without using the phrase "Venn diagram" or even clearly
stating how the lists of "specific" genes were derived. In all these cases,
the analysis boils down to comparing results significant in one, but not in
another comparison.

The use of Venn diagrams to illustrate specific differences between
comparisons should therefore be abandoned in favor of statistically correct
approaches. Furthermore, gene set enrichment analysis must never be applied
to sets of genes defined as significant in one comparison, but not the
other.

## Methods

### Methods availability

This document has been written as an Rmarkdown file, containing all
statistical calculations required to replicate the findings and figures.
The markdown file, along with additional files required to recreate this
manuscript have been uploaded to [https://github.com/bihealth/manuscript_venn_diagrams](https://github.com/bihealth/manuscript_venn_diagrams).

### Data

The expression data as a count matrix has been downloaded from GEO,
accession GSE156063.

### Statistical analyses

Power calculation was done using the R package pwr, version
`r packageVersion("pwr")`.
For differential gene expression, the R package DESeq2, version
`r packageVersion("DESeq2")`
has been used. Gene set enrichments were done using either hypergeometric
test (where stated) or the CERNO test using the package tmod [@zyla2019gene],
version
`r packageVersion("tmod")`.
GO terms have been sourced from the R package msigdbr, version
`r packageVersion("msigdbr")`.

### Literature survey

An informal literature survey was performed using Google Scholar to
estimate the frequency of the incorrect use of Venn diagrams. We searched
for articles from 2020 mentioning the phrases "differential expression" and "venn
diagram" in the Scientific Reports journal. The journal was
selected because it represents a wide spectrum of research areas, the
articles are available through an open access license and there is a large
number of publications per year. For each of the papers identified, we
checked whether (i) the authors used the VD to show differentially expressed
transcripts significant in one comparison, but not another, (ii) the
authors discussed "unique", "non-overlapping" or "specific" regions of the
Venn diagram and (iii) whether this was coupled to gene set enrichment
analysis is any form. Articles which (i) focused only on the intersections
of the Venn diagrams (genes common to all conditions), or (ii) which used the
Venn diagrams for a purpose other than to compare genes significant in one
condition, but not significant in another condition or (iii) for which a
clear-cut error could not be indicated past any reasonable doubt were not
considered incorrect.

## Bibliography