### 2022-04-05 mk5/manuscript.rmd
final update

---
title: "Venn diagrams may indicate erroneous statistical reasoning in transcriptomics."
author: "January Weiner^1,†^, Benedikt Obermayer^1^ and Dieter Beule^1^"
date: "`r Sys.Date()`"
outputoff:
  html_document
output:
  word_document:
    reference_docx: templates/template.docx
outputoff2: 
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
bibliography: manuscript.bib
link-citations: true
XXcsl: templates/nature-medicine.csl
---

^1^ Berlin Institute of Health at Charité – Universitätsmedizin Berlin, Core Unit Bioinformatics, Charitéplatz 1, 10117 Berlin, Germany

^†^ To whom the correspondence should be addressed

## Abstract

A common application of differential expression analysis is 
finding genes that are differentially expressed upon treatment in only one
out of several groups of samples. One of the approaches is to test 
for significant difference in expression between treatment and control separately
in the two groups, and then select genes that show statistical
significance in one group only. This approach is then often combined with
a gene set enrichment analysis to find pathways and gene sets
regulated by treatment in only this group.

Here we show that this procedure is statistically incorrect and that the
interaction between treatment and group should be tested instead. Moreover,
we show that gene set enrichment analysis applied to such incorrectly
defined genes group-specific genes may result in misleading artifacts.
Due to the presence of false negatives, genes significant in one, but not
the other group are enriched in gene sets which correspond to the
overall effect of the treatment.  Thus, the results appear related to the
problem at hand, but do not reflect the group-specific effect of a
treatment. A literature search revealed that more than a quarter of papers
which used a Venn diagram to illustrate the results of separate
differential analysis  have also applied this incorrect reasoning.

```{r,echo=FALSE}
## Set default options for the knitr RMD processing
knitr::opts_chunk$set(echo=FALSE,warning=FALSE,message=FALSE,fig.width=5,fig.height=5,cache=FALSE,autodep=TRUE, results="hide",dev="jpeg")
```

```{r libraries,cache=FALSE}
library(tidyverse)
library(limma)
library(GEOquery)
library(DESeq2)
## install bioshmods from github.com/bihealth/bioshmods
library(bioshmods)
library(tmod)
library(cowplot)
library(stringr)
library(ggvenn)
library(ggplot2)
library(ggpval)
library(eulerr)
library(Biobase)
library(pander)
library(org.Hs.eg.db)
library(pwr)
theme_set(theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()))
source("functions.R")
```

```{r setup}
padj_thr <- 0.05
lfc_thr  <- 1
n_stratum <- 10 # 2 strata (male and female) per group/treatment
                # total number of samples = n_stratum * 2 * 2 * 2
lfc_thresholds <- c(0, 0.5, 1, 1.5, 2, 2.5, 3)
```

## Introduction

Experimental designs for transcriptomic analyses frequently include more than one factor. Often, the
question asked is whether there is a difference between groups (first
factor) with respect to reaction to a particular treatment (second
factor). For example, we may ask whether there are differentially
expressed genes (DEGs) which are specific to a particular group of
patients, e.g. interferon response elicited by a virus in one group, but
absent in another group of patients. In other words, we ask whether the
difference between the control group (healthy subjects) and the treatment
group (infected patients) is different between two groups of individuals.
This "difference of differences" is known in statistics as an interaction
[@blalock1965theory]. To find out whether it is statistically significant,
an appropriate statistical test for interaction should be employed.

However, another approach is widely spread [@nieuwenhuis2011erroneous].
Instead of testing the interaction, the effect of the treatment is tested
separately in both groups. Next, a difference between groups is inferred if the 
effect of treatment
is significant in one comparison, but not significant in the other. This
approach is not correct from statistical point of view, as "the difference
between significant and not significant is not itself statistically
significant" [@gelman2006difference]. For example, the p-value in the first
comparison may be 0.009, and in the other comparison 0.011. At an
alpha level of 0.01 the difference will be statistically significant in the
first, but not significant in the other comparison.

In transcriptomics, statistical tests are performed for thousands of
genes. As in the general case, the inference of differences between the
groups should correctly be done by testing the significance of interaction
between the group and the treatment. In practice, the differences between
treatment and control are frequently tested in the two groups separately.
This can be visualized using a Venn diagram (VD, Fig. 1) showing the
overall number of DEGs significant in both comparisons (the intersection in
the VD) or significant in only one comparison (the remaining two fields on
a VD).  The genes which are significant in only one comparison are
sometimes incorrectly considered as specific for the corresponding group.
Following this, gene set enrichment analysis may be used
in an attempt to test which pathways are specific to one, but not the other
group.

In this paper, we show that under reasonable assumptions this approach may
result in apparent enrichments even if there are no real statistically
significant differences
between the groups. To this end, we randomly split a cohort into two
groups, compared the treatment (viral infection) with controls in each of
the groups separately and then applied gene set enrichment to the sets of
genes significantly different in one, but not the other group.
Moreover, we show that the resulting gene set enrichments correspond to the
differential expression between treatment and control. Thus, the enriched
terms are relevant to the biological question at hand, yet while they do
reflect real processes linked to viral infection, they do not correspond to
the differences between the study groups. Finally, we use literature search to
show that this incorrect approach to study group-specific treatment effects
is not uncommon. In fact, while VDs are a useful visualization tool also in
transcriptomics, in more than quarter of the papers where VDs were used,
group-specific genes were defined as significant in one, but not other
groups, and in 19% of the papers a gene set enrichment was performed.

## Results

### Transcriptomic changes due to Sars-Cov-2 infection

Consider two group of patients, G1 and G2 (Tab. 1). Each group contains 
`r n_stratum * 4`
individuals. In both groups, there is an equal number of healthy
individuals (labeled "Ctrl" on figures below) 
or patients infected with Sars-Cov-2 (labeled "SC2"). Our aim is to understand
the differences between G1 and G2 in the response to infection. For
example, we ask which genes or pathways are specifically up-regulated by SC2
infection in G1 as compared to G2, and vice versa. In the following, we
used the data set GSE156063 [@mick2020upper] in two approaches (an
incorrect and the correct one) to arrive at opposite conclusions.

**Table 1** Overall design in the case study: transcriptomic changes due to
Sars-Cov-2 infection. The table shows number of patients in each
combination of study group / disease status.

```{r results="markdown"}
tab <- as.data.frame(matrix(2 * n_stratum, ncol=2, nrow=2))
tab <- data.frame(c("", "Disease status", ""), c("", "Sars-Cov-2 infection", "Another infection"), 
  c("Group 1 (G1)", tab[[1]]), c("Group 2 (G2)", tab[[2]]))

colnames(tab) <- c("", "", "Study group", "")
pander(tab)
```

```{r pheno_data_download,eval=FALSE}
## this sometimes fails. The RDS of the object is provided
Sys.setenv(VROOM_CONNECTION_SIZE=8*131072)

## with getGEO, we can only get the phenoData
geo <- getGEO("GSE156063")[[1]]
saveRDS(geo, file="GSE156063.geo")
```

```{r pheno_data_cleanup}
geo <- readRDS("GSE156063.geo")
covar <- as(phenoData(geo), "data.frame")

## remove columns with only one value
boring <- map_lgl(covar, ~ length(unique(.x)) == 1)
covar <- covar[ , !boring ]

## clean up
covar <- covar %>% 
  dplyr::rename(gender = "gender:ch1") %>%
  mutate(disease = gsub(" .*", "", .data[["disease state:ch1"]])) %>%
  mutate(label = description) %>%
  mutate(group = disease) %>%
  arrange(description) %>%
  dplyr::select(all_of(c("title", "label", "gender", "disease", "group")))
```

```{r featuredata_download}
## the counts must be downloaded from GEO separately.
if(!file.exists("GSE156063_swab_gene_counts.csv.gz")) {
  download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE156nnn/GSE156063/suppl/GSE156063_swab_gene_counts.csv.gz",
                "GSE156063_swab_gene_counts.csv.gz")
}

counts <- read_csv("GSE156063_swab_gene_counts.csv.gz")
.tmp <- counts[[1]]
counts <- as.matrix(counts[,-1])
rownames(counts) <- .tmp
#counts <- counts[ , covar$description ]
counts <- counts[ , covar$label ]
lcpm   <- edgeR::cpm(counts, log=TRUE)
#stopifnot(all(colnames(counts) == covar$description))
stopifnot(all(colnames(counts) == covar$label))

annot <- data.frame(ENSEMBL = rownames(counts))

.tmp <- mapIds(org.Hs.eg.db, annot$ENSEMBL, column=c("ENTREZID"), keytype="ENSEMBL")
annot$ENTREZID <- .tmp[ match(annot$ENSEMBL, names(.tmp)) ]

.tmp <- mapIds(org.Hs.eg.db, annot$ENSEMBL, column=c("SYMBOL"), keytype="ENSEMBL")
annot$SYMBOL <- .tmp[ match(annot$ENSEMBL, names(.tmp)) ]

.tmp <- mapIds(org.Hs.eg.db, annot$ENSEMBL, column=c("GENENAME"), keytype="ENSEMBL")
annot$GENENAME <- .tmp[ match(annot$ENSEMBL, names(.tmp)) ]
```

```{r Preparation_of_the_covariates,cache=FALSE}
sel <- covar$group %in% c("no", "SC2")
counts <- counts[ , sel ]
covar  <- covar[ sel, ]
lcpm   <- lcpm[ , sel ]
covar$group <- as.character(covar$group)

set.seed(0123)

#covar$disease <- covar$group

g1 <- covar %>% mutate(n=1:n()) %>%
  group_by(gender, disease) %>% slice_sample(n=n_stratum) %>% pull(n)
g2 <- covar %>% mutate(n=1:n()) %>% filter(!n %in% g1) %>%
  group_by(gender, disease) %>% slice_sample(n=n_stratum) %>% pull(n)

covar$group <- NA
covar$group[g1] <- "G1"
covar$group[g2] <- "G2"

covar$group.disease <- paste0(covar$group, '_', covar$disease)

```

```{r DESeq2,cache=FALSE}
## DESeq2 calculations
## manual cache, since the operation takes a long time
sel <- !is.na(covar$group)

ds2 <- DESeqDataSetFromMatrix(counts[,sel], colData=covar[sel, ], design=~ 0 + group.disease )

ds2_file <- "ds2_cached.rds"
if(TRUE || !file.exists(ds2_file)) {
  message("Running DESeq2")
  ds2 <- DESeq(ds2)
  saveRDS(ds2, file=ds2_file)
} else {
  message("Reading ds2 from manual cache")
  ds2 <- readRDS(ds2_file)
}
```

```{r DE_analysis}
library(tidyverse)
library(tmod)
res <- list()
res$g1 <- results(ds2, contrast=c(-1, 1, 0, 0))
res$g2 <- results(ds2, contrast=c(0, 0, -1, 1))

res <- map(res, ~ .x %>% as.data.frame() %>%
  rownames_to_column("ENSEMBL") %>% 
  left_join(annot, by="ENSEMBL") %>%
  mutate(DEG=!is.na(padj) & abs(log2FoldChange) > lfc_thr & padj < padj_thr))

res.merged <- merge(res$g1, res$g2, by=c("ENSEMBL", "SYMBOL", "ENTREZID", "GENENAME"), suffixes=c(".g1", ".g2"))
```

First, we have performed differential gene expression analysis for each of
the groups G1 and G2 separately using standard bioinformatic tools. For each comparison, we defined
DEGs as genes for which the 
false discovery rate (FDR) was lower than 
`r padj_thr`
and absolute log~2~ fold change (LFC) was higher than
`r lfc_thr`. There were
`r sum(res$g1$DEG)` 
DEGs in the G1 group, and
`r sum(res$g2$DEG)`
in the G2 group. In total,
`r sum(res.merged$DEG.g1 & res.merged$DEG.g2)` 
DEGs were common for G1 and G2, 
`r sum(res.merged$DEG.g1 & !res.merged$DEG.g2)` 
DEGs were significant in G1 only ("specific" for G1), and 
`r sum(!res.merged$DEG.g1 & res.merged$DEG.g2)` 
were significant in G2 only (see Fig. 1, panel *A*).
A naive interpretation of these results implies that there is a substantial
difference between these two groups of individuals, as evidenced by a small
overlap in commonly regulated genes. The majority of DEGs is
significant in one comparison only.

```{r GO_db_prepare,cache=TRUE}
library(msigdbr)
min_mod_size <- 10
max_mod_size <- 50
msig_go_bp <- msigdbr(subcategory="GO:BP")
mset <- makeTmodFromDataFrame(msig_go_bp, 
                              feature_col="human_gene_symbol", 
                              module_col="gs_exact_source", 
                              title_col="gs_name")
mset$MODULES$N <- map_int(mset$MODULES$ID, ~ length(mset$MODULES2GENES[[.x]]))
mset <- mset[ mset$MODULES$N <= max_mod_size & 
              mset$MODULES$N >= min_mod_size ]
```

```{r GSEA,cache=FALSE}
common <- res.merged %>% filter(DEG.g1 & DEG.g2) %>% pull(SYMBOL)
gsea_res <- map(res, ~ {
  fg <- .x %>% filter(padj < padj_thr & abs(log2FoldChange) > lfc_thr)
  fg <- setdiff(fg$SYMBOL, common) # only "specific" genes
  tmodHGtest(fg=fg, bg=.x$SYMBOL, mset=mset)
})
gsea_res_full <- map(res, ~ {
  fg <- .x %>% filter(padj < padj_thr & abs(log2FoldChange) > lfc_thr)
  fg <- setdiff(fg$SYMBOL, common) # only "specific" genes
  tmodHGtest(fg=fg, bg=.x$SYMBOL, mset=mset, qval = Inf)
})
message("G1 ", nrow(gsea_res$g1), " G2 ", nrow(gsea_res$g2))
gsea_res.merged <- merge(gsea_res$g1, gsea_res$g2, by=c("ID", "Title"), suffixes=c(".g1", ".g2"), all=T)
saveRDS(gsea_res.merged, file="gsea_res_merged.rds")
```

To understand which
pathways are up-regulated in each of the two groups, we used a standard generation I
gene set enrichment analysis — a hypergeometric test — on the DEGs in each
group. Gene sets for the gene set enrichment analysis were taken from the Gene Ontology (GO) database.
Gene sets with more than `r max_mod_size` or 
fewer than `r min_mod_size` genes were removed.
For each group, we have
selected only genes which are DEGs in that group, but not the other,
mimicking a naive approach for finding pathways regulated in one patient
group only.
Here, a similar picture
emerged. Overall,
`r nrow(gsea_res$g1)` gene sets were significantly enriched in G1, and
`r nrow(gsea_res$g2)` gene sets were significantly enriched in G2. 
Both the Venn diagram (Fig. 1, B) and the results of
enrichments (Fig. 1, C and D) suggest that there is a
fundamental difference between the groups, and that the groups have little
in common in their response to the virus.

Importantly, the different GO terms enriched in the two groups were related
to infection, and may tempt to speculate about the underlying
biological differences between these two groups. For example, the significance of
Toll like receptor 4 pathway in G1, but not G2; and, vice versa, significance of
response to interleukin 7 in G2, but not in G1 may be considered as
evidence of altered immune response to the virus in G2 as compared to G1.

**Fig. 1. Results of differential gene expression analysis
and gene set enrichment analysis using an incorrect approach.**
**A**, Venn diagram showing numbers of differentially expressed genes (DEG)
in each of the two groups, G1 and G2; **B**, Venn diagram showing numbers
of significantly enriched GO terms in each of the two groups; **C** results
of gene set enrichment analysis for genes "specific" to group G1; **D**, results of gene set
enrichment analysis for genes "specific" to group G2 (only top 10 terms are shown).

```{r fig1, fig.width=15,fig.height=5.5,dpi=300,dev="jpeg"}
p1 <- ggvenn(list(G1=res$g1 %>% filter(DEG) %>% pull(ENSEMBL),
                         G2=res$g2 %>% filter(DEG) %>% pull(ENSEMBL)),
             show_percentage=FALSE) +
      labs(title="Differentially expressed genes") +
      ylim(-1, 1.4)
p2 <- ggvenn(list(G1=gsea_res$g1 %>% pull(ID),
                         G2=gsea_res$g2 %>% pull(ID)),
              show_percentage=FALSE) +
      labs(title="Enriched GO terms") +
      ylim(-1, 1.4)

col1 <- plot_grid(p1, p2, labels="AUTO", nrow=2)

p3 <- plot_gsea(gsea_res$g1 %>% arrange(P.Value) %>% dplyr::slice(1:10)) + ggtitle("G1") +
  scale_x_discrete(labels=function(x) str_wrap(x, width=25))
#p3 <- plot_grid(p3, NULL, ncol=1, rel_heights=c(2, 1))
p4 <- plot_gsea(gsea_res$g2 %>% arrange(P.Value) %>% dplyr::slice(1:10)) + ggtitle("G2") +
  scale_x_discrete(labels=function(x) str_wrap(x, width=25))

plot_grid(plotlist = list(col1, p3, p4), labels=c('', 'C', 'D'), nrow=1)
```

However, the groups G1 and G2 were randomly sampled from the same data set.
In fact, repeated re-sampling always results in some genes being found to
be significantly different in one group, but not the other, despite the
fact that one does not expect any major differences between sets of
individuals randomly drawn from a single population. Thus, the conclusions drawn
from a Venn diagram-driven gene set enrichment analysis are based on
artifacts. Closer
inspection of genes which are DEGs in one group, but not the other
reveals the underlying statistical fallacy (Fig. 2, *A* – *D*), that is, that difference
between significant and non-significant is, in itself, not statistically
significant [@gelman2006difference]. This does not necessarily mean that
there are no differences at all between these two groups, but that lack of
significance in one group and significance in the other group does not
correctly identify differences between groups.

```{r interaction}
res$int <- results(ds2, contrast=c(-1, 1, 1, -1)) %>%
  as.data.frame() %>%
  rownames_to_column("ENSEMBL") %>% 
  left_join(annot, by="ENSEMBL") %>%
  mutate(DEG=!is.na(padj) & abs(log2FoldChange) > lfc_thr & padj < 0.05)

fg <- res$int %>% filter(padj < padj_thr & abs(log2FoldChange) > lfc_thr)
if(nrow(fg) > 0) {
  gsea_res$int <-  tmodHGtest(fg=fg$SYMBOL, bg=res$int$SYMBOL, mset=mset)
}

tmod_res <- list()
tmod_res$int <- tmodCERNOtest(res$int %>% arrange(pvalue) %>% pull(SYMBOL),
                              mset=mset)
```

```{r correlation_coefs}
## calculate correlation coefficients for various groups of genes
## all correlations are significant

cors <- list()
cors$all <- with(res.merged, 
                cor(log2FoldChange.g1, log2FoldChange.g2))
cors$sign <- with(res.merged %>% filter(DEG.g1 | DEG.g2), 
                cor(log2FoldChange.g1, log2FoldChange.g2))
cors$g1 <- with(res.merged %>% filter(DEG.g1 & !DEG.g2), 
                cor(log2FoldChange.g1, log2FoldChange.g2))
cors$g2 <- with(res.merged %>% filter(!DEG.g1 & DEG.g2), 
                cor(log2FoldChange.g1, log2FoldChange.g2))

```

To find genes which are differentially regulated in the two groups, the
correct statistical approach is to calculate interaction between groups
(G1, G2) and disease status (no disease vs. COVID). While it may be argued
that a test for interaction has lower power than a test for a simple
contrast, no 
genes show a significant interaction even at FDR < 0.1. In fact, this is not
surprising. The log~2~ fold changes for comparisons withing G1 and G2 are
strongly correlated (Fig. 2, *E*). For all significant genes, the Pearson correlation
coefficient is
`r format.pval(cors$sign, digits=2)`, 
while for genes
exclusively significant in G1 or G2 (genes "specific" to G1 or G2), it is
`r format.pval(cors$g1, digits=2)` 
and
`r format.pval(cors$g2, digits=2)`, 
respectively. Thus, genes which are significant in one, but not in the
other comparison tend to have have similar log~2~ fold changes in both
groups (e.g. Fig.
2, *A*, *C* and *D*).

**Fig. 2. Genes which are significant in one comparison, but not
the other do not show a statistically significant interaction.** **A** -
**D**, examples of genes which are DEG in one group, but are not
significantly different in the other group. "Ctrl", healthy individuals; 
"SC2", Sars-Cov-2 infected patients. Values above the plot indicate FDR (p-values
corrected for multiple testing). **E**, correlation between log~2~ fold
changes in G1 and G2. Color indicates genes which are are significant in
one, but not significant in the other comparison; red indicates genes significant in G1, 
blue indicates genes significant in G2. The overall Pearson correlation
coefficient between log~2~ fold changes is 
`r with(res.merged, format.pval(cor(log2FoldChange.g1, log2FoldChange.g2), digits=2))`.

```{r fig2, fig.width=10, fig.height=5.5, dpi=300,dev="jpeg"}
#id    <- gsub(" ", "_", toupper("response to interferon beta"))
#gstitle    <- "GOBP_TOLL_LIKE_RECEPTOR_4_SIGNALING_PATHWAY"
gstitle    <- "GOBP_INTERLEUKIN_7_MEDIATED_SIGNALING_PATHWAY"
gsid  <- gsea_res$g2 %>% filter(grepl(gstitle, Title)) %>% pull(ID)
gsgenes <- mset$MODULES2GENES[[gsid]]
sel   <- res.merged %>% filter(SYMBOL %in% gsgenes) %>% arrange(pvalue.g2) %>% 
  filter(DEG.g2 & !DEG.g1) %>%
  dplyr::slice(1:4) %>% pull(ENSEMBL)

sample_sel <- c(g1, g2)

group_lab <- gsub("_no", "/Ctrl", gsub("_SC2", "/SC2", covar$group.disease))

pees <- map(sel, ~ {
  s1 <- res$g1 %>% dplyr::slice(match(.x, ENSEMBL)) %>% pull(padj)
  s2 <- res$g2 %>% dplyr::slice(match(.x, ENSEMBL)) %>% pull(padj)
  id <- res$g1 %>% dplyr::slice(match(.x, ENSEMBL)) %>% pull(SYMBOL)

  ggplot_gene(lcpm[.x, sample_sel ], 
              group_lab[ sample_sel ],
              annotation=format.pval(c(s1, s2), digits=2), textsize=8,
              pval_text_adj=.5) +
            ggtitle(id)
})

part1 <- plot_grid(plotlist=pees, labels="AUTO", ncol=2)

tmp <- res.merged %>% mutate(specific= 
                             ifelse(DEG.g1 & !DEG.g2,
                                    "G1", 
                                    ifelse(!DEG.g1 & DEG.g2, "G2", "NS"))) %>%
                      mutate(tlr4=SYMBOL %in% gsgenes)

part2 <- ggplot(tmp, aes(x=log2FoldChange.g1, 
                                y=log2FoldChange.g2, 
                                color=specific)) + 
         scale_color_manual(values=c(G1="red", G2="blue", NS="#666666")) +
         geom_point(alpha=.5) + xlab("G1") + ylab("G2") +
         geom_hline(yintercept=0, color="grey") +
         geom_vline(xintercept=0, color="grey") +
         geom_abline(slope=1, intercept=0, color="grey") +
         guides(color=FALSE)

plot_grid(part1, part2, rel_widths=c(2, 2), labels=c("", "E"))
```

Consequently, it is not possible to calculate gene set enrichment for the
interaction using a
hypergeometric test, as there are no DEGs for the interaction contrast.
Gene set enrichment using a second generation algorithm (CERNO), relying on the
ordering of genes according to their raw p-values from the interaction
contrast rather than selecting a
set of DEGs [@zyla2019gene], does not show any significant enrichment.

### Artifacts arise because of false negatives

It is worth noting that in the gene set enrichment analysis of the genes
"specific" for a given comparison – i.e., genes which are significant in
that comparison, but not significant in others – we have observed a number
of terms associated with immune response. It is a crucial point of this
manuscript to note that the spurious enrichments not only show significant
p-values, but also that the terms or pathways which appear in them are
relevant to the research hypothesis being tested. Below, we will show why
these terms (rather than random terms which have no obvious relevance to an
infectious disease) appear in the results.

To understand how significant results appear in a gene set enrichment
analysis in randomly generated groups despite absence of genes with
significant interaction, it is first necessary to consider the definition
of a differentially expressed gene in this context. More often than not,
DEGs are defined by a threshold in p-value adjusted for multiple testing,
possibly combined with a threshold in log~2~ fold change. The commonly used
Benjamini-Hochberg procedure [@benjamini1995controlling] ensures that among
genes for which FDR < 0.05 there are at most 5% false
positives irrespective of the sample size.

This way, we can exert control over the false positive rate (FPR, type I
errors), keeping it at a relatively low level. However, we do not control
the false negative rate (FNR, type II errors).  In a powerful statistical
test (such as a t-test), the test power in a typical application will
rarely achieve more than 80%. For example, even for large
effects (Cohen's d > 0.8) and type I error rate of 0.05, a t-test only achieves
80% power with at least 
`r floor(pwr.t.test(power=.8, d=0.8)$n)`
samples per group. For small effects (Cohen's d > 0.2), the required number
of samples is at least
`r floor(pwr.t.test(power=.8, d=0.2)$n)`
per group.
Even assuming a test power of 80%, the FNR is 20%. Clearly, false negatives (FNs) occur
at much higher rates than false positives (FPs). In the case of high throughput
data sets, where the FPR is controlled by Bejnamini-Hochberg procedure or a
similar technique, the FNR may be even as high as 80% [@white2019beyond].

```{r full_dataset}
ds2_full <- DESeqDataSetFromMatrix(counts, colData=covar, design=~ disease )
ds2_full<- DESeq(ds2_full)
res$full <- results(ds2_full, name="disease_SC2_vs_no") %>% as.data.frame() %>%
  rownames_to_column("ENSEMBL") %>% 
  left_join(annot, by="ENSEMBL") %>%
  mutate(DEG=!is.na(padj) & abs(log2FoldChange) > lfc_thr & padj < padj_thr)
```

```{r g_specific}
g1_spec <- res.merged %>% filter(DEG.g1 & !DEG.g2) %>% pull(ENSEMBL)
g2_spec <- res.merged %>% filter(!DEG.g1 & DEG.g2) %>% pull(ENSEMBL)
```

These FNs occur
at a much higher rate within the sets of DEGs defined by the
non-overlapping areas of the VDs, that is DEGs considered to be "specific"
for one group or other in a naive approach.
To illustrate this phenomenon, we have analyzed the full data set from which G1 and G2
were drawn (Fig. 3), comparing the 
`r sum(covar$disease == "no") `
healthy controls to 
`r sum(covar$disease == "SC2") `
COVID-19 patients. Of the
`r length(g1_spec)`
genes significant in G1, but not in G2, 
`r .n1 <- res$full %>% filter(DEG & ENSEMBL %in% g1_spec) %>% nrow; .n1`
(`r sprintf("%.0f%%", .n1/length(g1_spec) * 100)`) 
are significant in the full data set; 
of the
`r length(g2_spec)`
genes significant in G2, but not in G1, 
`r .n2 <- res$full %>% filter(DEG & ENSEMBL %in% g2_spec) %>% nrow; .n2`
(`r sprintf("%.0f%%", .n2/length(g2_spec) * 100)`) 
are significant in the full data set.
Given that G1 and G2 were sampled from the total population, and since the
FDR was set to `r padj_thr`, we do not expect more than 
`r .nfp <- ceiling(padj_thr * sum(res$full$DEG)); .nfp`
FPs in the full data set, 
which implicates that at least 
`r .n1 + .n2 - .nfp` out of the `r length(g1_spec) + length(g2_spec)` 
"specific" DEGs are true positives in the full data set.
Thus, we can assume that at least a third of
the genes that appeared to be "specific" in the initial analysis were, in
fact, false negatives in one of the comparisons.

In other words, a substantial fraction of the "specific" genes are genes
that are in reality differentially expressed in both groups alike.
Therefore, if one is to perform a gene set enrichment analysis on one of
these "specific" groups of genes, then the enriched functions will be
related to the pathways and processes up- or down-regulated in both groups
due to the common factor (in this example, the COVID-19 disease), but
which are not related to differences between the two groups.

**Fig. 3. Area-proportional Venn diagram showing overlaps in DEGs between
G1, G2 and the full data set.** The majority of genes which have been labeled
as DEGs in only one of the groups G1, G2 are DEGs when all data were
analyzed.

```{r fig3,dpi=300,dev="jpeg"}
tmp <- merge(res.merged, res$full, by="ENSEMBL")
tmp <- tmp[ , c("DEG.g1", "DEG.g2", "DEG") ]
colnames(tmp) <- c("G1", "G2", "Full dataset")
plot(euler(tmp), quantities=list(type=c("counts", "percent")))
```

### Influence of sample size and cut-off thresholds on number of artifacts

In the example above, the groups have been randomly sampled from a larger
data set only once. Arguably, the observations might differ if the groups
were to be resampled. Furthermore, we have chosen a group size of 
`r n_stratum * 4` (`r n_stratum * 2` per group/treatment combination). 
Larger sample sizes are known to increase robustness of gene set enrichment
analysis, and group size of 20 has been shown to be relatively robust [@maleki2019size].
However, a smaller or larger group size might change the proportion of FNs in the
results and thus influence the results of gene set enrichment. Finally,
we have used a log~2~ fold change threshold of `r lfc_thr`, because raising
it would increase the fraction of FNs. However, filtering for biologically
relevant genes with a substantially higher effect size may influence
the observed enrichments. On the other hand, raising the log~2~ fold change
threshold or lowering the sample size may lead to a smaller number of
defined DEGs, thus making the hypergeometric test less powerful and lead to
fewer gene set enrichment.

To investigate whether the sampling had an impact on
the artifacts, we have repeated the above procedure for 100 replicates,
each containing a different set of samples randomly assigned to the two
groups.  Furthermore, we tested whether the selection of the log~2~ fold
change threshold or different sample size might have an impact on the
extent of the arising artifacts. To this end, we have tested 
`r length(lfc_thresholds)` 
different threshold values for the log~2~ fold
changes and three sample sizes: 20, 40 or 80 samples per group
(corresponding to 10, 20 or 40 samples per group/treatment combination).

**Fig. 4. Influence of log~2~ fold change treshold (A, C) and the sample
size (B, D) on the number of genes significant in one, but not in the other
group (A, B) and the number of significant gene sets found in one group, but
not the other (C, D).**  Top row shows the influence of log~2~ fold change
threshold for sample size of 40 per group (20 per group/treatment
combination), bottom row shows the influence of sample size for a log~2~
fold change threshold set to `r lfc_thr`.

```{r montecarlo,cache=TRUE,fig.width=8,fig.height=6}
## note: given that the monte carlo simulation is quite time consuming, we
## have included the code for it in another file, `montecarlo.R`. If trying
## to replicate this manuscript, please execute that file first to generate
## the necessary replicates.

outdir <- "montecarlo"
merged_res_fn <- file.path(outdir, "merged_res.rds")

if(!file.exists(merged_res_fn)) {
  warning("Running simulations. This will take a long time.")
  source("montecarlo.R")
}

merged_res <- readRDS(file.path(outdir, "merged_res.rds"))

lfc_thresholds <- c(0, 0.5, 1, 1.5, 2, 2.5, 3)

fdr_thresholds <- c(.8, .7, .6, .5, .4, .3, .2, .1, .05, .01, .001, .0001, 0.00001, 1e-6)
by_fdr <- map_dfc(fdr_thresholds, ~ {
                      a <- get_specific_num(merged_res[["10"]], lfc_thr=0, padj_thr=.x)
                      b <- get_total_num(merged_res[["10"]], lfc_thr=0, padj_thr=.x)
                      (b - a)/b }
)
colnames(by_fdr) <- fdr_thresholds
by_fdr <- by_fdr %>% pivot_longer(everything(), names_to = "lfc")
by_fdr$lfc <- factor(by_fdr$lfc, levels=fdr_thresholds)
g0 <- ggplot(by_fdr, aes(x=lfc, y=value)) + geom_violin() +
  ylab("DEGs") + xlab("FDR threshold")

by_lfc_thr <- map_dfc(lfc_thresholds, ~ 
                      get_specific_num(merged_res[["10"]], lfc_thr=.x,
                                       padj_thr=padj_thr))
colnames(by_lfc_thr) <- lfc_thresholds
by_lfc_thr <- by_lfc_thr %>% pivot_longer(everything(), names_to = "lfc")

g1 <- ggplot(by_lfc_thr, aes(x=lfc, y=value)) + geom_boxplot() +
  ylab("DEGs") + 
  xlab(expression(log[2]~fold~change~threshold)) +
  theme(text=element_text(size=16))

gsea_files <- list.files(path=outdir, pattern="gsea_.*rds", full.names=TRUE)
gsea_by_lfc_thr_files <- gsea_files[ grepl("gsea_res_10", gsea_files) ]
names(gsea_by_lfc_thr_files) <- gsub("montecarlo/gsea_res_10_([0-9.]*)\\.rds", "\\1",
                                     gsea_by_lfc_thr_files)
gsea_by_lfc_thr_files <- gsea_by_lfc_thr_files[ order(as.numeric(names(gsea_by_lfc_thr_files))) ]
gsea_by_lfc_thr <- map_dfc(gsea_by_lfc_thr_files, gsea_specific_num)
gsea_by_lfc_thr <- gsea_by_lfc_thr %>% pivot_longer(everything(), names_to = "lfc")

g2 <- ggplot(gsea_by_lfc_thr, aes(x=lfc, y=value)) + geom_boxplot() +
  ylab("Enriched gene sets") + 
  xlab(expression(log[2]~fold~change~threshold)) +
  theme(text=element_text(size=16))

by_ssize <- map_dfc(c("5", "10", "20"), 
                    ~ get_specific_num(merged_res[[.x]], lfc_thr=1,
                                       padj_thr=padj_thr))
colnames(by_ssize) <- c(10, 20, 40)
by_ssize <- by_ssize %>% pivot_longer(everything(), names_to = "ss")

g3 <- ggplot(by_ssize, aes(x=ss, y=value)) + geom_boxplot() +
  ylab("DEGs") + xlab("Samples per group/treatment") +
  theme(text=element_text(size=16))

gsea_by_ss_files <- gsea_files[ grepl("gsea_res_[0-9]*_1\\.rds", gsea_files) ]
names(gsea_by_ss_files) <- gsub("montecarlo/gsea_res_([0-9]*)_.*\\.rds", "\\1",
                                     gsea_by_ss_files)
gsea_by_ss_files <- gsea_by_ss_files[ order(as.numeric(names(gsea_by_ss_files))) ]
gsea_by_ss <- imap(gsea_by_ss_files, ~ {
                     data.frame(ss=.y, value= gsea_specific_num(.x))
                                     })
gsea_by_ss <- Reduce(rbind, gsea_by_ss)
gsea_by_ss$ss <- factor(as.numeric(gsea_by_ss$ss) * 2, levels=c(10, 20, 40))

g4 <- ggplot(gsea_by_ss, aes(x=ss, y=value)) + geom_boxplot() +
  ylab("Enriched gene sets") + xlab("Samples per group/treatment") +
  theme(text=element_text(size=16))

plot_grid(g1, g2, g3, g4, labels = LETTERS[1:4])
```

Setting a higher log~2~ fold change threshold reduces the number of
DEGs as well as of the observed artifactual gene set enrichments 
(Fig. 4, A and B). However, even for log~2~
threshold of 3 (DEGs defined by 8-fold change and FDR < `r padj_thr`) the
number of replicates in which artifacts can be observed is 
`r sum(gsea_by_lfc_thr %>% filter(lfc == 3) %>% pull(value) > 0)`
(out of 100),
and in at least 
`r sum(gsea_by_lfc_thr %>% filter(lfc == 3) %>% pull(value) >= 5)`
replicates, 5 or more gene sets were significantly
enriched. Thus, setting a more conservative threshold while retaining the
incorrect procedure cannot fully protect from the arising artifacts.
The number of artifacts rose with sample size
(Fig. 4, C and D). For total sample size of 160 (40 samples per
group/treatment combination) the number of artifacts was almost 
`r round(
  mean(gsea_by_ss %>% filter(ss == 40) %>% pull(value)) /
  mean(gsea_by_ss %>% filter(ss == 20) %>% pull(value)))
`
times higher than for total sample size of 80.

### Incorrect analysis of interactions is common in transcriptomics

@nieuwenhuis2011erroneous observed incorrect analyses of interactions in
about half of the analyzed papers from top neuroscience journals where the
authors considered an experimental design allowing for a test for
interaction. We wanted to know if this problem is
common in transcriptomics, too. To this end, we have searched three
journals – from broad to specialized – for the occurrence of the terms
"differential expression" with "venn diagrams" (Tab. 2). Next, we analyzed
the selected papers from year 2020 (and years 2015-2020 in one case) to
decide whether the VD was described or referred to as showing genes
"specific" or "unique" to a particular group or whether a test for
interaction was performed. Finally, we checked whether gene set enrichment
analysis was applied to genes significant in one, but not the other
group in order to find group-specific differences.

```{r lit_surveys}
get_keywords <- function(file, pattern="^([a-zA-Z]+): *(.*)") {
  ll <- readLines(file)
  ll <- ll[ grepl(pattern, ll) & !grepl("^https://", ll) ]

  ret <- gsub(pattern, "\\2", ll)
  names(ret) <- gsub(pattern, "\\1", ll)
  ret <- as.list(ret)
  ret$file <- file
  ret$incorrect_count <- get_incorrect_counts(file)
  ret$incorrect_count_vdonly <- get_incorrect_counts(file, vdonly=TRUE)
  ret
}

get_incorrect_counts <- function(file, vdonly=FALSE) {

  ll <- readLines(file)

  pat <- "^https://"
  if(vdonly) {
    sum(grepl(pat, ll) & grepl("vd only", ll))
  } else {
    sum(grepl(pat, ll))
  }

}

files <- list.files(pattern="literature_survey.*.md")

finfo <- map(files, get_keywords) 
names(finfo) <- map_chr(finfo, ~ .x[["journal"]])
```

**Tab. 2** Results of the informal literature survey. 
We searched for papers using Google Scholar and the keywords "venn diagram"
and "differential expression".
*Journal*, journal title; *Years*, publication dates; *Total*, total number
of papers found using the search phrase; *Analyzed*, total number of papers
which have been analyzed for correctness; *Incorrect*, total number
(percentage) of
papers in which the Venn diagram was combined with comparing significance
to non-significance; *Enrichment*, total number (percentage) of papers
which combined the Venn diagram with a gene set enrichment analysis.

```{r results="markdown"}
tab <- map_dfr(finfo, ~ 
               data.frame(Journal=.x$journal,
                          Years=.x$years,
                          Total=.x$total,
                          Analyzed=.x$analysed,
                          Incorrect=.x$incorrect_count,
                          Enrichment=.x$incorrect_count - .x$incorrect_count_vdonly))  %>%
  mutate(Analyzed=as.numeric(Analyzed)) %>%
  mutate(Total=as.numeric(Total))

tab <- rbind(tab, 
             data.frame(Journal="**Total:**", Years="", Total=sum(tab$Total),
                        Analyzed=sum(tab$Analyzed),
                        Incorrect=sum(tab$Incorrect),
                        Enrichment=sum(tab$Enrichment)))
tab2 <- tab %>%
  mutate(Incorrect=sprintf("%d (%.0f%%)", Incorrect, 100 * Incorrect / Analyzed)) %>%
  mutate(Enrichment=sprintf("%d (%.0f%%)", Enrichment, 100 * Enrichment / Analyzed))

pander(tab2)
```

We found 
that of the 
`r tab2$Analyzed[nrow(tab2)]` analyzed articles
 which used the terms "venn diagram" and
"differential expression", at least 
`r tab2$Incorrect[nrow(tab2)]`
were using Venn diagrams to compare
statistical significance with lack thereof by referring to "unique",
"specific", "solely regulated" or "exclusive" DEGs. Out of these, at least 
`r tab$Enrichment[nrow(tab2)]`
coupled the VDs with some form of gene set enrichment analysis on the
set of supposedly "specific" DEGs. In summary, in at least a quarter of the
papers on differential expression in which a Venn diagram was used, it was
illustrating an incorrect statistical procedure which may result in
artifactual gene set enrichments.

## Discussion

Drawing conclusions from comparing significance with lack thereof is a
common statistical fallacy [@gelman2006difference]. Just as absence of evidence is not evidence of
absence, the failure to reject the null hypothesis does not constitute the
same level of evidence as rejecting it. However, when such an incorrect
analysis is combined with downstream functional analysis, the
resulting pathways or gene ontologies are misleadingly relevant. For
example, the identified gene sets are associated with immune response for
a research hypotheses involving an infectious disease, or cancer pathways if
the underlying research hypothesis involved cancer treatment. Such results
may appear reasonable in the given context, especially if the correct
analysis of interactions does not show any significant differences. This
effect is persistent or even exacerbated for larger sample sizes (Fig.
4).

We found that this type of incorrect analysis occurs in more than a quarter
of papers where the procedure was illustrated with a VD.  That is not to
say that VDs are not a useful tool, even in the context of transcriptomics
and gene set enrichments, if used correctly. For example, gene set
enrichment analysis of an intersection of DEGs (i.e., by considering genes
from the overlap in a VD) is not an incorrect procedure.  Genes in the
overlapping part of a VD are significant in both (or all) comparisons,
hence no comparison between significance and non-significance is made.

While VDs appear to be frequently associated with an incorrect statistical
reasoning, the use of VDs is not the cause.  In the absence of a VD
illustrating the DEGs common and unique to the different study groups, two
incorrect approaches may still be found. Firstly, the direct
comparison of gene set enrichment results: that is, drawing conclusions
from the fact that a gene set enrichment result was significant in one
comparison only.  Second, while VDs are often used to illustrate
the numbers of "specific" DEGs and so present a mean to find examples of
this fallacy in scientific literature, researches test for enrichment
these "specific" genes without using the phrase "Venn diagram" or even
clearly stating how the lists of "specific" genes were derived. In all
these cases, the analysis boils down to comparing results significant in
one, but not in another comparison.

As an alternative to Venn diagrams and the downstream gene set enrichment
analysis, two approaches can be considered.
The correct statistical approach, as shown above, involves a test for interaction which
can reveal genes for which the impact of treatment significantly differ
between the groups. The results can then be plugged into a gene set
enrichment analysis the usual way.  Unfortunately, this has two major
drawbacks. Firstly, effect sizes (log~2~ fold changes) of the interaction
term are harder to interpret than log~2~ fold changes in a direct, group vs
group comparison.  The effect size in an interaction is negative if the
log~2~ fold change in the first comparison is larger than the log~2~ fold
change in the second comparison; this is, however, irrespective of whether
the differences in the individual comparisons are negative or positive,
which makes it harder to separate the differences in genes up-regulated in
one or both groups.

The second problem may arise if the changes are similar in both
comparisons, but of larger magnitude in one of them. For example, in a time
series context, the changes may be more pronounced at a later time point.
In this case, the analysis will show that the processes enriched for the
interaction term are the same as those enriched in each of the comparisons
individually.  While the results of the gene set enrichment analysis in
this context are correct, the result may not be what the researchers
intended – processes which qualitatively (rather than quantitatively)
differ between the comparisons.

An alternative approach, discordance/concordance analysis, has been proposed by @domaszewska2017concordant,
aiming at identifying processes which qualitatively differ between the two
comparisons. Here, a heuristic score ("disco score") has been defined which depends on the
effect sizes and p-values in both comparisons. The sign of the score
depends on whether the effects have the same sign (concordant; genes
up-regulated in both comparisons or down-regulated in both comparisons) or
opposite signs (discordant; genes up-regulated in one, but down-regulated
in the other comparison, and genes down-regulated in one, but up-regulated
in the other comparison). While the score does not allow the calculation of
a p-value and does not present an alternative to an analysis of
interaction, it can facilitate both visualization and further analysis
using a gene set enrichment algorithm.

Visualization of interaction for individual genes is straightforward (see
for example Fig. 2, *A-D*). However, the point of VDs is to show a grand
overview of the whole analysis summarizing thousands of results for the
analyzed genes. As an alternative of such an overview, we suggest plotting
the log~2 fold changes in one comparison against log~2~ fold changes in the
second comparison. This allows an intuitive assessment of the differences
between the two comparisons, in especially in combination with color coding
the genes which either are significant in the interaction or by coloring
using the disco score (see Fig. 2, *E* for an example).

Defining group-specific genes based on significant difference in one,
but no significant difference in another comparison is thus more than only a
statistical fallacy leading to erroneous results. When combined with gene
set enrichment analysis it can lead to potentially
sound-looking, and therefore particularly misleading results. This method
of obtaining specific differences between groups should therefore be
abandoned in favor of statistically correct approaches.  Furthermore, gene
set enrichment analysis must never be applied to sets of genes defined as
significant in one comparison, but not the other.

## Methods

### Methods availability

This manuscript has been written in R markdown [@xie2018r].  
All statistical calculations required to replicate the findings and figures
are contained in the source R markdown file.  The R markdown file, along with
additional files required to recreate this manuscript as well as the
results
of literature survey have been uploaded to
[https://github.com/bihealth/manuscript_venn_diagrams](https://github.com/bihealth/manuscript_venn_diagrams).

### Data

The expression data as a count matrix has been downloaded from GEO,
accession GSE156063.

### Statistical analyses

Power calculation was done using the R package pwr, version
`r packageVersion("pwr")`.
For differential gene expression, the R package DESeq2, version
`r packageVersion("DESeq2")`
has been used. Gene set enrichments were done using either hypergeometric
test (where stated) or the CERNO test using the package tmod [@zyla2019gene],
version
`r packageVersion("tmod")`.
GO terms have been sourced from the R package msigdbr, version
`r packageVersion("msigdbr")`.

### Simulation study

We have generated replicates of the example study for different log~2~ fold
change thresholds (`r paste(lfc_thresholds, sep=", ")`) and three different
sample sizes (40, 80 and 160, corresponding to sample size per
group/treatment combination of 10, 20 and 40). For each replicate, the full
procedure as described above was repeated, and numbers of DEGs and
significantly enriched terms were collected.

### Literature survey

A literature survey was performed using Google Scholar to
estimate the frequency of the incorrect use of Venn diagrams. We searched
for articles including the phrases "differential expression" and "venn
diagram" in three journals: Scientific Reports (2020), Nature Communications (2020) and
Science Immunology (2015-2020). For each of the papers identified, we
checked whether (i) the authors used the VD to show differentially expressed
transcripts significant in one comparison, but not another, (ii) the
authors discussed "unique", "non-overlapping" or "specific" regions of the
Venn diagram and (iii) whether this was coupled to gene set enrichment
analysis is any form. Articles which (i) focused only on the intersections
of the Venn diagrams (genes common to all groups), or (ii) which used the
Venn diagrams for a purpose other than to compare genes significant in one
groups, but not significant in other groups or (iii) for which a
clear-cut error could not be identified past any reasonable doubt were not
considered incorrect.
The results of the literature survey (including links to papers classified
as incorrectly analysing the interaction) are included in the manuscript
sources.

```{r}
sessioninfo::session_info(to_file=TRUE)
```

## Bibliography